A Kochen-Specker system has at least 21 vertices

### 2014-04-14 main.tex
Some more work.

% TODO this suggests only one result: the improved lower bound
%      reference to other results?
\title{A Kochen-Specker system has at least 21 vertices}

@@

Kochen and Specker proved that such a theory cannot satisfy:
% TODO note contextuality
\begin{spin}
    Given three pairwise orthogonal directions.
    In exactly one of the directions, the particle will not deviate.
\end{spin}
Their argument is based on the existence of a Kochen-Specker system.
\begin{dfn}
    A \keyword{Kochen-Specker (KS) system} is
    a finite set of points on the sphere
    for which each pair is not antipodal and
    there is no~\keyword{010-coloring}.
    A $010$-coloring is a~$\{0,1\}$-coloring of the points such that
    \begin{enumerate}
        \item
            no pair of orthogonal points are both colored~$1$ and
        \item
            of three pairwise orthogonal points exactly one is colored~$1$;
            or alternatively: they are colored~$0$, $1$ and~$0$ in some order.
    \end{enumerate}
\end{dfn}
% TODO shorten this proof.
A point on the sphere obviously corresponds to a direction in space.
Because of this, the term point, vector and direction
can be used interchangeably. Antipodal points correspond to oppositve
vectors and these span the same direction in space.

@@

The KS system proposed by Kochen and Specker contained 117 points\cite{ks}.
% TODO reference (x2)
Penrose and Peres independently found a smaller system of 33 points.
The current record is the 31 point system of Conway.
As pointed out by \cite{c00,aow11}, finding small KS systems
is of both theoretical and practical interest.
In public lectures, Conway himself, stressed the search for small KS systems.

@@

First, in Subsection~\ref{sec:ks33} we consider the 33 vector KS system
of Penrose and Peres.  Then, in Subsection~\ref{sec:ksgraphs},
we repeat a part of AOW's work, in particular the reduction of
KS systems to graphs.
The bottleneck of their search was the sheer number of graphs
and the deciding embeddability.

In Section~\ref{sec:ilb},
we improve upon their reduction,
to cut down the number of graphs drastically,
and state the results of our main computation.
Finally, in Section~\ref{sec:emb},
we describe our practical embeddability test.

\subsection{The 33 vector KS system of Penrose and Peres}
\label{sec:ks33}
TODO

@@

The requirements can be mechanically converted
to a formal sentence
in the language of the real numbers.
This sentence is true if and only if the graph is embeddable.
Using variations on Tarski's quantifier elimination procedure,
we can decide the truth of these sentences.
We have used the redlog\cite{redlog} package of the reduce algebra
system.  In Appendix~\ref{adx:unemb-10-2} the Reader can find
the reduce script generated mechanically for the graph
in Figure~\ref{fig:unemb-10-2}.
% TODO references to quantifier elimination
%

\section{Future research}
TODO

@@

% attribute mckay for answering questions?
\bibliography{main}{}
\bibliographystyle{plain}

>>>>

\title{A Kochen-Specker system has at least 21 vertices}

@@

Kochen and Specker proved that such a theory cannot satisfy:
% TODO note contextuality
\begin{spin}
    Given three pairwise orthogonal directions.
    In exactly one of the directions, the particle will not deviate.
\end{spin}
Their argument is based on the existence of a Kochen-Specker system.
\begin{dfn}
    A \keyword{Kochen-Specker (KS) system} is
    a finite set of points on the sphere
    for which each pair is not antipodal and
    there is no~\keyword{010-coloring}.
    A $010$-coloring is a~$\{0,1\}$-coloring of the points such that
        \footnote{
                In other papers, like \cite{aow11},
                the~$0$ and~$1$ are swapped:
                101-colorings.
                Then the colors correspond directly to the squared measurement
                outcome of the spin.}
    \begin{enumerate}
        \item
            no pair of orthogonal points are both colored~$1$ and
        \item
            of three pairwise orthogonal points exactly one is colored~$1$;
            or alternatively: they are colored~$0$, $1$ and~$0$ in some order.
    \end{enumerate}
\end{dfn}
% TODO shorten this proof.
A point on the sphere obviously corresponds to a direction in space.
Because of this, the term point, vector and direction
can be used interchangeably. Antipodal points correspond to oppositve
vectors and these span the same direction in space.

@@

The KS system proposed by Kochen and Specker contained 117 points\cite{ks}.
Penrose and Peres\cite{peres} independently found a smaller system of 33 points.
The current record is the 31 point system of Conway\cite[p.~197]{qtcm}.
As pointed out by \cite{c00,aow11}, finding small KS systems
is of both theoretical and practical interest.
In public lectures, Conway himself, stressed the search for small KS systems.

@@

First, in Subsection~\ref{sec:ksgraphs},
we repeat a part of AOW's work, in particular the reduction of
KS systems to graphs.
The bottleneck of their search was the sheer number of graphs
and the deciding embeddability.

In Section~\ref{sec:ilb},
we improve upon their reduction,
to cut down the number of graphs to consider drastically,
and state the results of our main computation.
Finally, in Section~\ref{sec:emb},
we describe our practical embeddability test.

@@

The requirements can be mechanically converted
to a formal sentence
in the language of the real numbers.
This sentence is true if and only if the graph is embeddable.
Famously, Tarski proved\cite{tarski}
that such sentences are decidable.
His decision procedure has an impractical complexity.
However, its practical value has been improved
by, for instance, the method of cylindrical algebraic decomposition\cite{qecad}.
We have used the redlog\cite{redlog} package of the reduce algebra
system, which implements a variant of Tarski's quantifier elimination.
In Appendix~\ref{adx:unemb-10-2} the Reader can find
the reduce script generated mechanically for the graph
in Figure~\ref{fig:unemb-10-2}.

\section{Conclusion and future research}
TODO

@@

% attribute mckay for answering questions?
% TODO attribute geng
\bibliography{main}{}
\bibliographystyle{plain}
